---
title: "Practical Machine Learning Assignment"
author: "IcedMcstuffin"
date: "2026-04-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction
In this project, my goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website [here](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).

```{r message = FALSE}
library(dplyr)
library(caret)
```

# 2. Data Cleaning
Data Cleaning has been performed by collecting all the columns with accelerometer reading (20) and removing the var_accel parameters giving us 16 accelerometer predictors for 4 accelerometers on x,y,z axis and the total accelerometer reading
```{r echo = TRUE}
## DATA CLEANING
# Reading the data and cleaning it to have only accelerometer columns and classe column and setting classe to factor class
rawtraining <- read.csv("pml-training.csv")
acceltraining <- select(rawtraining, contains(c("accel", "classe")))
acceltraining$classe <- as.factor(acceltraining$classe)
training <- acceltraining[,-c(2,7,12,17)]
training$classe <- as.factor(training$classe)

rawtesting <- read.csv("pml-testing.csv")
acceltesting <- select(rawtesting, contains("accel"))
testing <- acceltesting[,-c(2,7,12,17)]
```

# 3. Model Selection
Since there are 5 possible outcomes, we will be checking the random forest predictor model, gradient boosting model and assessing each of its accuracy

For all the test, a 5 fold cross validation is performed by trainControl() function as the training data set has large number of osbervations - 19622.

## 3.1 Random Forest model

```{r cache=TRUE}
set.seed(3223)
rfmodel <- train(classe~., data=training, method="rf", trControl = trainControl(method = "cv", number = 5))
print(rfmodel)
```
This shows that the expected out-of-sample accuracy of this model is 0.95 and the error rate is 5%

This is the plot for the Variable importance of each of the predictor.
```{r}
plot(varImp(rfmodel))
```

## 3.2 Gradient Boosting model

```{r cache=TRUE}
set.seed(3223)
gbmModel <- train(classe~., data=training, method="gbm", trControl = trainControl(method = "cv", number = 5), verbose = FALSE)
print(gbmModel)
```
This shows that the expected out-of-sample accuracy of this model is 82% and the error rate is 12%

# 4. Conclusion
After comparing the accuracy of these models, we will choose to select the model because of its out of sample accuracy of 95%







